---
title: "30 Day Map Challenge"
author: "Wei Jing Ang"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    theme: journal
    toc_float: true
    code_folding: hide
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup_packages, warning = FALSE, message = FALSE}
# Load libraries
library(tidyverse)
library(sf)
library(ggplot2)
library(patchwork)
library(ghibli)
library(ggrepel)
library(ggtext)
library(glue)
library(tigris)
library(rgeoboundaries)
library(raster)
library(classInt)
library(terra)
library(biscale)
library(cowplot)
library(viridis)
library(rayshader)
library(magick)
library(ggpubr)
```

```{r load_data, message = FALSE, warning = FALSE, results = 'hide'}
# Mekong river basin
basin <- read_sf("Data/Mekong_basin_delta/Mekong_basin_delta.shp")
countries_mekong <- read_sf("Data/Mekong_basin_countries/Mekong_basin_countries.shp")
countries <- read_sf("Data/World_Countries_(Generalized)/World_Countries__Generalized_.shp")

# Dams
dams <- read_sf("Data/Dams_1055/Dams_1055.shp")
dams1980 <- read_sf("Data/1980/1980.shp") # 1980
dams1990 <- read_sf("Data/1990/1990_combi.shp") # 1990
dams2000 <- read_sf("Data/2000/2000_combi.shp") # 2000
dams2010 <- read_sf("Data/2010/2010_combi.shp") # 2010
dams2020 <- read_sf("Data/2020/2020_combi.shp") # 2020

# River and lake
river1 <- read_sf("Data/River_1/FLOW1_STRA_3_China.shp") # upper course
river2 <- read_sf("Data/River_2/FLOW1_STRA_5.shp") # middle course
river3 <- read_sf("Data/River_3/FLOW3_CLASS2_TS.shp") # lower course
lake <- read_sf("Data/Lake/HydroLAKES_TS_finer.shp") # Tonle Sap Lake

# Amazon river
amazon_river <- st_read("Data/Amazon_hydrorivers/Amazon_hydrorivers.shp")
amazon_basin <- st_read("Data/Amazon_hydrobasin/Amazon_hydrobasin.shp")
curuai <- st_read("Data/CR/CR.shp") # shapefile for single lake

# Bathymetry
bathy <- rast("Data/Bathymetry/bathymetry.tif") # SpatRaster (rater() gives RasterLayer)
bathy_values <- values(bathy)[!is.na(values(bathy))] # remove na for classification
breaks <- classInt::classIntervals(bathy_values, 
                                   style = "quantile", n = 4)
bathy_class <- terra::classify(bathy, breaks$brks) # split by quantiles
bathy_df <- as.data.frame(bathy_class,xy=TRUE) # convert to df for mapping

# River polygon
hydropoly <- read_sf("Data/Hydropoly/hydropoly_recropped.shp")
hydropoly_add <- read_sf("Data/Hydropoly_additional/Hydropoly_additional.shp")

# USA
usa_wateruse_state <- read_csv("Data/usa_wateruse_state.csv")
usa_wateruse_penn <- st_read("Data/usa_wateruse_penn/usa_wateruse_penn.shp")
usa_states <- states(cb = TRUE)
hexbin <- read_sf("Data/us_states_hexgrid.geojson")
```

# Overall theme: Water

# 1. Points
Dams felt like a good fit for day 1's theme on points, and I wanted to convey the message that the many dams being planned and under construction in the Mekong River Basin will expand the network to over a thousand in the next decade. 

To draw attention to the upcoming dams, I employed the preattentive attributes of intensity and color. The current dams are in grey and of lower opacity, while the upcoming dams are in orange and of higher opacity. The Mekong River Basin is in dark grey to create a clear visual hierarchy in contrast with the light grey basemap. 

Using Gestalt's principle of similarity, part of the annotation is colored according to the orange new dams. To highlight that the dam network is growing exponentially, "over a thousand" was bolded, following the preattentive attribute of line width. 

```{r day1, message = FALSE, warning = FALSE, fig.align = "center",fig.height=7,fig.width=4.5,out.width='60%'}
# Define annotation 
annotation1 <- glue("<span style='color:#C66125;'>**Dams planned and under<br>construction**</span> are expected to<br>expand the network to **over a<br>thousand** in the next decade.")

ggplot() +
  
  # Map limits
  xlim(93, 110) +
  ylim(9.2, 33.2) +
  
  # Basemap (countries)
  geom_sf(data=countries, fill = "grey92", col = "#FCF8F2", lwd = 0.9)+
  
  # Mekong basin
  geom_sf(data = basin, fill = "grey30", color = "transparent") + # lower basin
  geom_sf(data = rbind(river1, river2, river3), color = "#3981A4", lwd=0.9) + # river
  geom_sf(data = lake, fill = "#3981A4", color = "transparent") + # Tonle Sap Lake

  # Operating dams
  geom_sf(data = dams %>% 
            filter(Status == "Operational"),
          pch = 21, aes(fill = "grey"), color = "transparent", size = 2) +

  # Planned/ constructing dams
  geom_sf(data = dams %>% 
            filter(Status == "Under construction" | Status == "Planned"),
          pch = 21, aes(fill = "#C66125"), color = "transparent", size = 2) +
  
  # Define colors
  scale_fill_manual(values = c(alpha("#C66125",0.8),alpha("grey",0.5)))+
  
  # Labels
  geom_text(aes(x = 109.5, y = 31.2, label = "01 Points"), size = 13/.pt, hjust=1) +
  
  geom_text(aes(x = 93, y = 10.5, label = "Wei Jing Ang\n30 Day Map Challenge"), hjust = 0, size = 5/.pt) +
  
  labs(title = "Dams in the Mekong",
       caption = "Source: Ang, W.J., Park, E., Pokhrel, Y., Tran, D.D., 
Loc, H.H., 2023, Replication Data for: Dams in the Mekong:
A comprehensive database, spatiotemporal distribution, and 
hydropower potentials, https://doi.org/10.21979/N9/ACZIJN.")+
  
  # Annotations
  geom_richtext(aes(x = 102.5, y = 29,
                    label = annotation1),
                size = 9/.pt, 
                colour = "grey30",
                hjust = 0,
                fill = NA,
                label.color = NA)+
  
  # Define map theme
  theme_void() +
  theme(plot.title = element_text(face="bold",size=15,margin=margin(0,20,-40,0),hjust=1),
        plot.caption = element_text(size=5, hjust = 0,margin=margin(-30,0,0,14)),
        # make sure to include color=NA to not have a thin grey border!
        panel.background = element_rect(fill = "#FCF8F2", color = NA),
        legend.position = "none")

ggsave("Output/01-WeiJing-Points.png", plot = last_plot(), height = 7, width = 4.5,units = "in")
```

# 2. Lines
I chose to map rivers for day 2's theme on lines. Focusing on the Amazon River Basin, I used the preattentive attribute of color. The river basin is in black to create a clear visual hierarchy in contrast with the light beige background, and to make the lines representing the rivers stand out. The rivers are colored and sized according to their stream order, where the largest streams of the highest orders are brighter and thicker.

```{r day2,fig.height=6,fig.width=6, fig.align='center',out.width='80%'}
# Project data to SAD 1969 Lambert South America
amazon_river <- amazon_river %>%
  st_transform('ESRI:102015')
amazon_basin <- amazon_basin %>%
  st_transform('ESRI:102015')

ggplot()+
  
  # Basin
  geom_sf(data=amazon_basin,fill="black",col=NA)+ 
  
  # Rivers by stream order
  geom_sf(data=amazon_river, aes(col=ORD_STRA,linewidth=ORD_STRA)) + 
  scale_color_viridis(option="A",begin=0.3)+
  scale_linewidth(range = c(0.5, 1.5)) + 
  
  # Labels
  labs(title = "The Amazon River Basin",
       subtitle = "02 Lines",
       caption = "Wei Jing Ang\n30 Day Map Challenge\n\nSource: HydroSHEDS")+
  
  # Theme
  theme_void()+
  theme(legend.position = "none",
        plot.title = element_text(size=16,face="bold",margin=margin(20,0,-20,20)),
        plot.subtitle = element_text(size=14,margin=margin(30,0,-30,20)),
        plot.background = element_rect(fill="#FCF8F2", color = NA),
        plot.caption = element_text(size=7,margin=margin(-25,15,15,0)))

#ggsave("Output/02-WeiJing-Lines.png", plot = last_plot(), height = 6, width = 6,units = "in")

## Code used for data wrangling ----
# hydrorivers <- st_read("Data/HydroRIVERS_v10_sa_shp/HydroRIVERS_v10_sa_shp/HydroRIVERS_v10_sa.shp")
# amazon_river <- hydrorivers %>%
#   filter(MAIN_RIV == "60443230") %>%
#   filter(ORD_STRA>4)
# st_write(amazon_river, "Amazon_hydrorivers/Amazon_hydrorivers.shp")
# 
# hydrobasins <- st_read("Data/hybas_sa_lev01-06_v1c/hybas_sa_lev03_v1c.shp")
# amazon_basin <- hydrobasins %>% filter(MAIN_BAS=="6030007000")
# st_write(amazon_basin, "Amazon_hydrobasin.shp")
```

# 4. Hexagons
For the theme on hexagons, I wanted to map the US states by water use. In line with the preattentive attribute of intensity, states with higher water usage are in dark blue, while states with lower usage are in lighter blue. The spherical mercator projection was used to create hexagons of equal sizes. 

```{r day4, fig.height = 5, fig.width = 7, out.width = '100%'}
# Join to hexagons (total withdrawals in million gallons per day (Mgal/d))
usa_wateruse_hex <- left_join(hexbin %>%
                                dplyr::select("iso3166_2","geometry"),
                              usa_wateruse_state,
                              by = c("iso3166_2"="STATE")) %>%
  st_transform('EPSG:3857') # Spherical Mercator for equal hexagon sizes


# Categorize data (convert to billion gallons)
usa_wateruse_hex$bin <- cut(usa_wateruse_hex$Use,
  breaks = c(0,2000, 5000, 10000, 20000,Inf),
  labels = c("0-2",">2-5", ">5-10", ">10-20",">20-28"),
  include.lowest = TRUE
)

# Define palette
palette_water <- c("#90e0ef","#00b4d8","#0188BC","#015BA0","#03045e")

ggplot(data=usa_wateruse_hex)+
  
  # Hexagons and state labels
  geom_sf(aes(fill=bin), color = "transparent", alpha = 0.9) +
  geom_sf_text(aes(label = iso3166_2), color = "white", size = 3, alpha = 0.6) +
  
  # Labels
  labs(title = "Water withdrawals by state in 2015",
       subtitle = "04 Hexagons",
       caption = "Wei Jing Ang\n30 Day Map Challenge\n\nSource: Dieter, C.A., Linsey, K.S., Caldwell, R.R., Harris, M.A.,
Ivahnenko, T.I., Lovelace, J.K., Maupin, M.A., Barber, N.L., 2018,
Estimated Use of Water in the United States County-Level Data for 2015,
https://doi.org/10.5066/F7TB15V5. ANDREWXHILL, us_states_hexgrid.") +
  
  # Colours
  scale_fill_manual(values = palette_water,
                    name = "Billion gallons per day",
                    guide = guide_legend(keyheight = unit(3, units = "mm"),
                                         keywidth = unit(10, units = "mm"),
                                         label.position = "bottom", title.position = "top", nrow = 1))+
  
  # Map theme
  theme_void() +
  theme(legend.position = "inside",
        legend.position.inside = c(0.565, 0.865),
        legend.text = element_text(size=9),
        legend.title = element_text(size=9),
        plot.background = element_rect(fill = "#FCF8F2", color = NA),
        panel.background = element_rect(fill = "#FCF8F2", color = NA),
        legend.background = element_rect(fill = "#FCF8F2", color = NA),
        plot.caption = element_text(size=5,hjust=1, 
                                    margin = margin(b = 0.5, t = -0.7, r = 0.5, unit = "cm")),
        plot.title = element_text(size = 16,face="bold",
                                  margin = margin(b = -0.7, t = 0.6, l = 3.5, unit = "cm")),
        plot.subtitle = element_text(size = 13,
                                  margin = margin(b = -1.6, t = 1.4, l = 3.5, unit = "cm")))

#ggsave("Output/04-WeiJing-Hexagons.png", plot = last_plot(), height = 5, width = 7,units = "in")

## Code used for data wrangling ----
# usa_wateruse <- read_csv("Data/Water_Use/usco2015v2.0.csv",skip=1)
# Sum water usage per state
# usa_wateruse_state <- usa_wateruse %>%
#   dplyr::select("STATE","TO-Wtotl") %>%
#   rename(Water = "TO-Wtotl") %>%
#   group_by(STATE) %>%
#   summarise(Use = sum(Water, na.rm = T))
# write.csv(usa_wateruse_state,"usa_wateruse_state.csv", row.names = FALSE)
```

# 6. Raster
This map of the Amazon River's floodplain lakes was derived from satellite imagery, flooding frequency, and field surveys. It is a continuous surface data, representing the topography (depth) of the floodplain lakes along the middle-lower Amazon. 

To highlight the lakes, I employed the preattentive attribute of color. The lakes are in brighter red, while the river is in darker blue. As the data is continuous, the lakes are colored as a gradient and classified as quantiles for easier interpretation of the legend. 

Following Gestalt's principle of similarity, parts of the annotation are colored according to the blue Amazon River and red lakes. To highlight that the Amazon floods frequently, "7 months" was bolded, employing the preattentive attribute of line width.

```{r day6, message=FALSE,warning=FALSE,out.width='100%', fig.width=8, fig.height=3.5}
# Define annotation 
annotation6 <- glue("<span style='color:#355B7C;'>**The Amazon**</span> floods for up to **7 months**<br>each year, forming <span style='color:#DD6F78;'>**large complex lakes<br>of varying depths**</span> along its banks.")

# Define palette
amazon_palette = c("#F1A790","#DD6F78","#BF6476","#8E6781")

ggplot()+
  
  # Map limits
  xlim(-60, -54) +
  ylim(-3.8, -1.5) +
  
  # Lake depths
  geom_tile(data = bathy_df, aes(x = x, y = y, fill = bathymetry), color = NA) + 
  
  # River
  geom_sf(data=hydropoly, fill = "#355B7C", color = NA) + 
  geom_sf(data=hydropoly_add, fill = "#355B7C", color = NA) + 
  
  # Define colors
  scale_fill_manual(values=amazon_palette,
                    name = "Depth (m)",
                    labels = c("2","","","10"),
                    guide=guide_legend(direction="horizontal",title.vjust = 0.8))+
  
  # Labels
  labs(title = "The Amazon River's floodplain lakes",
       subtitle = "06 Raster",
       caption="Wei Jing Ang\n30 Day Map Challenge\n\nSource: Ang, W.J., Edward, P., Enner, A., 2021, 
Related data for: Mapping floodplain bathymetry in the 
middle-lower Amazon River using inundation frequency 
and field control, https://doi.org/10.21979/N9/CEPE6Y.")+

  # Annotations
  geom_richtext(aes(x = -60, y = -2.2,
                    label = annotation6),
                size = 11/.pt,
                colour = "grey30",
                hjust = 0,
                fill = NA,
                label.color = NA)+
  
  # Define map theme
  theme_void()+
  theme(plot.title = element_text(face="bold",size=14,margin=margin(20,0,-30,30)),
        plot.subtitle = element_text(size=12,margin=margin(40,0,-40,30)),
        plot.caption = element_text(size=5, hjust = 1,margin=margin(-40,10,10,0)),
        plot.background = element_rect(fill = "#FCF8F2", color = NA),
        panel.background = element_rect(fill="#FCF8F2",color=NA),
        legend.text = element_text(size=9),
        legend.title = element_text(size=9),
        legend.position="inside",
        legend.position.inside = c(0.5,0.07),
        legend.title.position = "left",
        legend.text.position = "bottom",
        legend.key.height = unit(0.5, "cm"),
        legend.key.width=unit(0.5, "cm"))

#ggsave("Output/06-WeiJing-Raster.png", plot = last_plot(), height = 3.5, width = 8,units = "in")
```

# 7. Vintage style
The vintage map is a remake of day 1's map on the Mekong dam network. As the prompt asked for a map with a historical feel but on a contemporary topic, I thought that mapping modern day hydropower dams would be a good fit.  

To draw attention to the upcoming dams, I used the preattentive attributes of intensity and size. The current dams are the smaller crosses of lower opacity, while the upcoming dams are the bigger crosses of higher opacity. The Mekong River Basin is in dark brown to create a clear visual hierarchy in contrast with the light brown basemap. 

The colour choices are more muted to match the theme. The vintage background was obtained from pinterest. Since unlike day 1, size instead of color was used to differentiate the dams, I added a legend. 

```{r day7, message = FALSE, warning = FALSE, fig.align = "center",fig.height=7,fig.width=4.5,out.width='60%'}
# Define annotation 
annotation7 <- glue("Dams planned and under<br>construction are expected to<br>expand the network to over a<br>thousand in the next decade.")

# Load background image (from pinterest)
vintage <- image_read("Data/Vintage_bg.jpg")

ggplot() +
  
  background_image(vintage) +
  
  # Map limits
  xlim(93, 110) +
  ylim(9.2, 33.2) +
  
  # Basemap (countries)
  geom_sf(data=countries, fill = "transparent", col = alpha("#3a2317",0.4), lwd = 0.3)+
  
  # Mekong basin
  geom_sf(data = basin, fill = "#9C846B", color = "transparent") + # lower basin
  geom_sf(data = rbind(river1, river2, river3), color = "#C4C6AE", lwd=0.9) + # river
  geom_sf(data = lake, fill = "#C4C6AE", color = "transparent") + # Tonle Sap Lake

  # Operating dams
  geom_sf(data = dams %>% 
            filter(Status == "Operational"),
          pch = 4, aes(size = 1), color = alpha("#3a2317",0.4)) +

  # Planned/ constructing dams
  geom_sf(data = dams %>% 
            filter(Status == "Under construction" | Status == "Planned"),
          pch = 4, color = "#3a2317", aes(size = 2)) +
  
  # Define sizes
  scale_size_continuous(range=c(1,2),
                        name = "",
                        breaks=c(1,2),
                        labels = c("Operational","Planned and\nunder construction"))+
  
  # Labels
  geom_text(aes(x = 109.5, y = 31.2, label = "07 Vintage"), size = 13/.pt, hjust=1) +
  
  geom_text(aes(x = 93, y = 11, label = "Wei Jing Ang\n30 Day Map Challenge"), hjust = 0, size = 5/.pt) +
  
  labs(title = "Dams in the Mekong",
       caption = "Source: Ang, W.J., Park, E., Pokhrel, Y., Tran, D.D., 
Loc, H.H., 2023, Replication Data for: Dams in the Mekong:
A comprehensive database, spatiotemporal distribution, and 
hydropower potentials, https://doi.org/10.21979/N9/ACZIJN.
Vintage background was obtained from pinterest.")+
  
  # Annotations
  geom_richtext(aes(x = 102.5, y = 29,
                    label = annotation7),
                size = 9/.pt, 
                colour = "grey30",
                hjust = 0,
                fill = NA,
                label.color = NA)+
  
  # Define map theme
  theme_void() +
  theme(plot.title = element_text(face="bold",size=15,margin=margin(0,20,-38,0),hjust=1),
        plot.caption = element_text(size=5, hjust = 0,margin=margin(-40,0,0,14)),
        legend.position = "inside",
        legend.position.inside = c(0.78, 0.7),
        legend.text = element_text(size=8),
        legend.background = element_rect(fill = NA, color = NA))

#ggsave("Output/07-WeiJing-Vintage.png", plot = last_plot(), height = 7, width = 4.5,units = "in")
```

# 12. Time and space
For a map that visualizes change over space and time, I thought that adapting one of my previous maps on hydropower growth patterns over time in the Mekong River Basin would be a good match.

I employed the preattentive attribute of intensity, where countries with lower hydropower capacity are in light purple, while the countries with higher capacity are in dark purple. Using Gestalt's principle of similarity, part of the title is colored according to Laos in orange, to highlight the growing hydropower in the country.

```{r day12, out.width='100%', fig.width=7,fig.height=4}
# Define function to sum hydropower per country per decade
bycountry <- function(hpyear,year){
  hpyear %>% 
  st_drop_geometry() %>%
  dplyr::select(Capacity_M, Country) %>%
  group_by(Country) %>%
  summarise(Capacity = sum(Capacity_M, na.rm = T)) %>%
  mutate(decade = year)
}

# Run function to sum hydropower
# Note: Myanmar only appears in 2010, so add in a row for Myanmar until 2000
country1980 <- bycountry(dams1980,1980) %>%
  rbind(.,data.frame(Country = "Myanmar",Capacity = 0, decade = 1980)) 
country1990 <- bycountry(dams1990,1990) %>%
  rbind(.,data.frame(Country = "Myanmar",Capacity = 0, decade = 1990))
country2000 <- bycountry(dams2000,2000) %>%
  rbind(.,data.frame(Country = "Myanmar",Capacity = 0, decade = 2000))
country2010 <- bycountry(dams2010,2010)
country2020 <- bycountry(dams2020,2020)
countryall <- rbind(country1980,country1990,country2000,country2010,country2020)

# Add geometry to countryall
countries_mekong <- countries_mekong %>% rename(Country = COUNTRY) # rename for joining
countryall_withgeom <- left_join(countryall,
                                 countries_mekong %>% 
                                   dplyr::select(geometry,Country),
                               by = "Country") %>% # left join geometries to country data
  st_as_sf(crs = st_crs(basin)) # turn into sf for mapping

# Decadal labels for the map
annotations <- data.frame(
  decade = c(1980, 1990, 2000, 2010, 2020),
  label = c("1980s", "1990s", "2000s", "2010s", "Post-2020s")
)

# Define title and subtitle 
title12 <- glue("Growing hydropower in <span style='color:#C66125;'>Laos</span> in the Mekong Basin")

# Maps of hydropower capacity over the decades
ggplot()+
  
  # Mekong delta
  geom_sf(data = basin, fill = "#D6D6FF", color = "transparent") + 
  
  # Capacity
  geom_sf(data = countryall_withgeom, aes(fill = as.factor(ntile(Capacity, 5))), # quintiles
          color = "transparent") +
  facet_wrap(~decade, nrow = 1) + # 1 map for each decade
  scale_fill_manual(values = c("#D6D6FF","#B2B4F0","#6B7ED4","#4561BF","#003995"),
                    name = "Gigawatts",
                    labels = c("0","","","","33"),
                    guide=guide_legend(direction="horizontal",title.vjust = 0.8))+
  
  # Laos
  geom_sf(data = countries_mekong %>%
            filter(NAME == "Laos"), fill = "transparent", color = "#C66125", linewidth = 1) +
  
  # Labels
  geom_text(data = annotations, aes(x = 100, y = 8, label = label), size = 9/.pt) + # decadal labels
  labs(title = title12,
       subtitle = "12 Time and space",
  caption = "Wei Jing Ang\n30 Day Map Challenge\n\nSource: Ang, W.J., Park, E., Pokhrel, Y., Tran, D.D., Loc, H.H., 2023, Replication 
Data for: Dams in the Mekong: A comprehensive database, spatiotemporal 
distribution, and hydropower potentials, https://doi.org/10.21979/N9/ACZIJN.") +
  
  # Define map theme
  theme_void() + 
  theme(
    strip.text.x = element_blank(), # remove facet wrap labels
    legend.position = "inside",
    legend.position.inside = c(0.18,-0.18),
    legend.title.position = "left",
    legend.text.position = "bottom",
    legend.key.height = unit(0.4, "cm"), # reduce legend key size
    legend.key.width = unit(0.4, "cm"),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8),
    plot.margin = margin(t = 5, r = 5, b = 5, l = 5),
    plot.background = element_rect(fill = "#FCF8F2", color = NA),
    plot.caption = element_text(hjust = 1, size=5, margin = margin(9,6,6,0)),
    plot.subtitle = element_text(size=12,margin=margin(10,0,10,7)),
    plot.title = element_markdown(size=12,face="bold", margin=margin(10,0,0,7))
  )

#ggsave("Output/12-WeiJing-SpaceTime.png", plot = last_plot(), height = 4, width = 7,units = "in")
```

# 16. Choropleth
Adapting the bivariate map we made in class, I mapped the water usage in Pennsylvania with the same dataset used on day 4, comparing it against average household sizes. Notable areas are identified and labelled. 

To focus on Pennsylvania, I employed the preattentive attribute of color. The vibrant colors create a clear visual hierarchy in contrast with the light grey basemap and beige background (representing water in this case). Following Gestalt's principle of similarity, parts of the title are colored according to the two variables.

```{r day16, out.width='100%',fig.width=7,fig.height=5.5}
# Rename columns and project data
usa_wateruse_penn <- usa_wateruse_penn %>%
  rename(Water = "PS.Wtotl") %>%
  rename(Household = "AVE_HH_SZ") %>%
  st_transform('EPSG:2272')

# Classify data into bivariate categories using 'biscale'
bivariate <- bi_class(usa_wateruse_penn, x = Water, y = Household, style = "quantile", dim = 3)

# Basemap
usa_states_pennproj <- usa_states %>%
  st_transform('EPSG:2272')

# Define title and annotations using colored text
title16 <- glue("<span style='color:#6EB4B5;'>**Water usage**</span> and <span style='color:#A25F99;'>**household size**</span> in Pennsylvania")
annotation16_1 <- glue("High water usage<br>small households")
annotation16_2 <- glue("High water usage<br>large households")
annotation16_3 <- glue("Low water usage, smaller households")

# Define palette
palette <- "DkBlue2"

# Create map with colored subtitle, annotations, and leader lines
map <- ggplot() +
  theme_void(base_size = 14) +
  
  # Plot basemap
  geom_sf(data=usa_states_pennproj, fill = "grey92", col = "white", lwd = 0.9)+
  
  # Plot bivariate map
  geom_sf(data = bivariate, aes(fill = bi_class), color = NA, linewidth = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = palette, dim = 4, flip_axes = FALSE, rotate_pal = FALSE)+
  
  # Plot outline
  geom_sf(data=st_union(bivariate),fill=NA,col="white",lwd=0.8)+
  
  # Set map limits
  coord_sf(xlim = c(1189585-120000, 2814853+100000), 
           ylim = c(140908.2-320000, 1075997+200000), expand = FALSE)+
  
  # Labels
  labs(title = title16,
       subtitle = "16 Choropleth",
       caption = "Wei Jing Ang\n30 Day Map Challenge\n\nSource: Dieter, C.A., Linsey, K.S., Caldwell, R.R., Harris, M.A., 
Ivahnenko, T.I., Lovelace, J.K., Maupin, M.A., Barber, N.L., 2018, 
Estimated Use of Water in the United States County-Level Data for 2015,
https://doi.org/10.5066/F7TB15V5. Esri Data and Maps, USA Counties.

Water usage refers to the total withdrawals from public supply per day in 2015") +
  
  # Define map theme
  theme(plot.title = element_markdown(size=16, face="bold",margin=margin(10,0,-5,50)),
        plot.subtitle = element_text(size=14, margin = margin(20,80,-75,0),hjust=1),
        plot.caption = element_text(size=6,hjust=0, margin = margin(-70,0,5,40)),
        panel.grid = element_blank(),
        panel.background = element_rect(fill="#FCF8F2",color=NA))+
  
  # Annotations
  geom_richtext(aes(x = 1220585, y = 310000,
                    label = annotation16_1),
                size = 11/.pt,
                colour = "white",
                hjust = 0,
                fill = NA,
                label.color = NA)+
  geom_richtext(aes(x = 2319085, y = 280000,
                    label = annotation16_2),
                size = 11/.pt,
                colour = "white",
                hjust = 0,
                fill = NA,
                label.color = NA)+
  geom_richtext(aes(x = 1870585, y = 900000,
                    label = annotation16_3),
                size = 11/.pt,
                colour = "white",
                hjust = 0,
                fill = NA,
                label.color = NA)

# Legend
legend <- bi_legend(pal = palette,   
                    flip_axes = FALSE,
                    rotate_pal = FALSE,
                    dim = 4,
                    xlab = "Water usage",
                    ylab = "Household size",
                    size = 7)+
  # Remove white legend background
  theme(plot.background = element_rect(fill = NA, color = NA))

# Combine map and legend using cowplot
finalPlot <- ggdraw() +
  draw_plot(map, 0, 0, 1, 1) +
  draw_plot(legend, 0.7, 0.02, 0.21, 0.21)

# Display final map
finalPlot

ggsave("Output/16-WeiJing-Choropleth.png", plot = last_plot(), height = 5.5, width = 7,units = "in")

## Code used for data wrangling ----
# usa_wateruse <- read_csv("Data/Water_Use/usco2015v2.0.csv",skip=1)
# counties <- st_read("Data/USA_Counties_1016676858958179337.geojson")
# usa_wateruse_penn <- left_join(counties %>%
#                                  filter(STATE_FIPS == "42") %>%
#                                  dplyr::select("FIPS","STATE_FIPS","AVE_HH_SZ","geometry"),
#                                usa_wateruse %>%
#                                  dplyr::select("FIPS","COUNTY","STATE","PS-Wtotl"),
#                                by = "FIPS") 
# st_write(usa_wateruse_penn, "usa_wateruse_penn.shp")
```

# 18. 3D
For the theme on 3D, mapping the lake depths from day 6 felt like a good fit. I zoomed into one prominent lake and exaggerated the depths for them to show up in the 3D plot, created using the rayshader package. 

```{r day18, fig.height=3.5,fig.width=7,out.width='100%'}
# Crop bathymetry to a single lake (Curuai)
bathy_cropped <- crop(bathy, curuai)
bathy_masked <- mask(bathy_cropped, curuai)

# Convert to matrix to use rayshader package
matrix = raster_to_matrix(bathy_masked)
matrix = 0-matrix*800 # exaggerate depth and convert to negative elevation

# Add shadow and water to 3D plot
matrix %>%
  sphere_shade(sunangle = 40, zscale = 200, texture = "imhof2") %>%
  add_shadow(ray_shade(matrix, zscale = 200), 0.5) %>% #,
  plot_3d(matrix, zscale = 500, fov = 0, theta = 10, zoom = 0.4, phi = 45, windowsize = c(1000, 400),water = TRUE, waterdepth = 0, wateralpha = 0.5, watercolor = "lightblue",
             waterlinecolor = "white", waterlinealpha = 0.5,background="#FCF8F2") 
Sys.sleep(0.2)
render_snapshot("Data/Curuai_3D.png",clear=TRUE) # save snapshot

# Load snapshot and convert to raster
img <- image_read("Data/Curuai_3D.png")
img_gg <- as.raster(img)

# Label using ggplot
ggplot() +
  
  # Snapshot (now a raster)
  annotation_raster(img_gg, xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf) +

  # Labels
  labs(title = "Curuai Lake in the Amazon",
       subtitle = "18 3D",
       caption="Wei Jing Ang\n30 Day Map Challenge\n\nSource: Ang, W.J., Edward, P.; Enner, A., 2021, Related data for: 
Mapping floodplain bathymetry in the middle-lower Amazon River using 
inundation frequency and field control, https://doi.org/10.21979/N9/CEPE6Y.")+

  # Define map theme
  theme_void()+
  theme(aspect.ratio = 2/5,
        plot.title = element_text(face="bold",size=14,margin=margin(20,20,-25,0),hjust=1),
        plot.subtitle = element_text(size=12,margin=margin(40,20,-40,0),hjust=1),
        plot.caption = element_text(size=6,hjust=0,margin=margin(-25,0,10,20)),
        plot.background = element_rect(fill="#FCF8F2",color=NA))

#ggsave("Output/18-WeiJing-3D.png", plot = last_plot(), height = 3.5, width = 7,units = "in")
```

# 24. Only circular shapes
Mapping Curuai Lake again, I converted the raster grid to an sf object to map them as circles representing the lake depth. The raster was resampled to a coarser resolution for the circles to be drawn bigger and clearer. Using the preattentive attribute of intensity, deeper parts of the lake are in dark purple while shallower parts are in light red.

```{r day24, fig.height=3.8,fig.width=7,out.width='100%'}
# Resample Curuai lake to coarser grid
new_res <- 0.01
new_ras <- rast(ext(bathy_masked), resolution = new_res, crs = crs(bathy_masked))
curuai_rs <- resample(bathy_masked, new_ras, method = "bilinear")

# Classify raster into quantiles
curuai_values <- values(curuai_rs)[!is.na(values(curuai_rs))] # remove na for classification
curuai_breaks <- classInt::classIntervals(curuai_values, 
                                   style = "quantile", n = 4)
curuai_class <- terra::classify(curuai_rs, curuai_breaks$brks) # split by quantiles

# Convert to data frame for mapping as raster
curuai_df <- as.data.frame(curuai_class,xy=TRUE)

# Turn data frame to sf object to map as circles
curuai_sf = st_as_sf(curuai_df, coords = c("x", "y"), 
                 crs = 4326, agr = "constant")

ggplot()+
  
  # Curuai lake depth
  geom_sf(data = curuai_sf, aes(color=bathymetry), size=1.2) + 
  
  # Define colors
  scale_color_manual(values=amazon_palette,
                    name = "Depth (m)",
                    labels = c("2","","","7"),
                    guide=guide_legend(direction="horizontal",title.vjust = 0.8))+
  
  # Labels
  labs(title = "Curuai Lake in the Amazon",
       subtitle = "24 Only circular shapes",
       caption="Wei Jing Ang\n30 Day Map Challenge\n\nSource: Ang, W.J., Edward, P.; Enner, A., 2021, Related data for: 
Mapping floodplain bathymetry in the middle-lower Amazon River using 
inundation frequency and field control, https://doi.org/10.21979/N9/CEPE6Y.")+

  # Define map theme
  theme_void()+
  theme(plot.title = element_text(face="bold",size=14,margin=margin(20,20,-20,0),hjust=1),
        plot.subtitle = element_text(size=12,margin=margin(30,20,-30,0),hjust=1),
        plot.caption = element_text(size=6,hjust=0,margin=margin(-25,0,10,20)),
        plot.background = element_rect(fill="#FCF8F2",color=NA),
        panel.background = element_rect(fill="#FCF8F2",color=NA),
        legend.text = element_text(size=11),
        legend.title = element_text(size=11),
        legend.position="inside",
        legend.position.inside = c(0.84,0.7),
        legend.title.position = "left",
        legend.text.position = "bottom",
        legend.key.height = unit(0.5, "cm"),
        legend.key.width=unit(0.5, "cm"))

#ggsave("Output/24-WeiJing-Circles.png", plot = last_plot(), height = 3.8, width = 7,units = "in")
```

# 26. Map projections
Since my overall theme is on water, I wanted to try mapping the Spilhaus Projection in R, a projection that combines all oceans into one. However, this projection is not directly available in R, so I obtained its coordinates (in mercator projection) from a code shared on Github. 

```{r day26, warning = FALSE,fig.height=6,fig.width=6,dpi=300,out.width='80%',fig.align='center'}
# The coordinates for the spilhaus projection was extracted from github
# https://github.com/rtlemos/spilhaus/tree/main
spilhaus_coord = read.csv("Data/spilhaus_coord.csv")

# Annotation data frame
annotation26 <- data.frame(
   x = c(4200000,1000000,-7500000),
   y = c(-6000000,1500000,4000000),
   label = c("Pacific\nOcean", "Indian\nOcean","Atlantic\nOcean")
)

# Map
ggplot(data=spilhaus_coord, aes(x=x, y=y)) +
  geom_raster(fill="#008080") +
  coord_equal() +
  
  # Labels
  geom_text(data=annotation26, aes(x=x, y=y, label=label),
           color="white", 
           size=4,
           fontface="bold")+
  labs(title = "Spilhaus Projection",
       subtitle = "26 Map projections",
       caption = "Wei Jing Ang\n30 Day Map Challenge\n\nSource: rtlemos, spilhaus") +
  
  # Theme
  theme_void()+
  theme(plot.background = element_rect(fill = "#FCF8F2", color = NA),
        plot.title = element_text(size=16,face="bold",
                                  margin = margin(b = -2.6, t = 1.5, l = 7.2, unit = "cm")),
        plot.subtitle = element_text(size=14,
                                  margin = margin(b = -2.6, t = 3, l = 7.2, unit = "cm")),
        plot.caption = element_text(size=8,hjust=0,
                                    margin = margin(b = 1.2, t = -2, l = 0.8, unit = "cm")))

#ggsave("Output/26-WeiJing-Projections.png", plot = last_plot(), height = 6, width = 6,units = "in")
```
